### 2024-10-07 MetadataS1/MetadataS1.tex
Add star abundance of breeding table 4 and minor change title table 5

\begingroup  
  \centering
\textbf{Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut}\\[1.5em]
 Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Gilles Gauthier\textsuperscript{4}, Éliane Duchesne\textsuperscript{1}, 
 Dominique Fauteux\textsuperscript{5}, Dominique Berteaux\textsuperscript{1}, Pierre Legagneux\textsuperscript{3,6}, Marie-Christine Cadieux \textsuperscript{3} and Joël Bêty\textsuperscript{1}\\[1.5em]
\textbf{Author Affiliations}:\\
\textsuperscript{1} Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada\\
\textsuperscript{2} Chaire de Recherche du Canada en Écologie Intégrative, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université de Sherbrooke, Sherbrooke, QC, Canada\\
\textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’Impact des Migrations Animales sur les Écosystèmes Nordiques, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{4} Centre d’Études Nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{5} Centre d’Études Nordiques, Centre de Connaissance et d’Exploration de l’Arctique, Musée Canadien de la Nature, Ottawa, ON, Canada\\
\textsuperscript{6} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France 
\\[1.5em]
\textbf{Corresponding Author}:\\
\textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
\textit{\href{mailto:joel_bety@uqar.ca}{joel{\textunderscore}bety@uqar.ca}}
\\[1.5em]
\textbf{Open Research statement}:\\
The data set is publicly available at \url{https://datadryad.org/} and the raw data and codes used to extract the data set are publicly available at \url{https://zenodo.org/}.\\
\endgroup
\newpage

@@

Since 2004, brown and collared lemmings were live-trapped 3 times during the summer (mid-June, mid-July, and mid-August) in two 11 ha grids. Each grid is made of 144 traps separated by 30 m according to a cartesian plane, one in mesic habitat and the other in wet habitat, located in the Qarlikturvik valley \citep{fauteux2015, gauthier2020lemmings}. Density of each species was estimated at each occasion using spatially explicit capture-recapture methods (see \citet{fauteux2015} for details). From 1995 to 2016 snap-trapping was performed once a year (mid-July) along 2 groups of transects located in the same habitats than the trapping grids \citep{gruyer2008}. Index of abundance derived from snap-trapping were transformed in density estimates in each habitat for the period 1995-2003 using the equation provided by \citet{fauteux2018} based on the period of overlap between the two sampling methods (2004 to 2016). 

@@

All vertebrate individuals captured for marking purposes were systematically weighed: snow goose (G. Gauthier, M.-C. Cadieux and J. Lefebvre, unpublished data), snowy owl \citep{therrien2012, robillard2018}, American-golden plovers \citep{lamarre2021}, common-ringed plovers \citep{leandri2019}, other shorebirds (J. Bêty, unpublished data), glaucous gulls \citep{gauthier2015}, long-tailed jaeger \citep{seyer2019}, parasitic jaegers (J.-F. Therrien and G. Gauthier, unpublished data), Lapland longspurs (J. Bêty and G. Gauthier, unpublished data), lemmings \citep{gauthier2020lemmings}, ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}. When not available, we extracted mean body mass from the literature \citep{wilman2014}.

@@

    \subsection*{3.  Research methods}
                \subsubsection*{a. Field/laboratory}
We estimated the abundance of breeding individuals for most species, but there were a few exceptions. For common ravens, parasitic jaegers, long-tailed ducks, and king eiders, we suspect the presence of a significant number of non-breeding individuals in the study area. Therefore, the estimates we provided for these species include both breeding and potentially non-breeding individuals. Additionally, we did not distinguish between breeding and non-breeding individuals for mammals such as brown and collared lemmings, Arctic fox, ermine, and Arctic hare. The methods used for each species are summarized in (\textbf{Table \ref{table:table_summary_methods_and_abundance}}).

@@

                \item[] \textit{\textbf{King eider and long-tailed duck}}\\
                We estimated the abundance of both king eiders and long-tailed ducks based on the annual nest density of each species found in the 8 km\textsuperscript{2} extensive nest search area located in the Qarlikturvik valley. We extrapolated the mean nest density in the wetlands of the Qarlikturvik valley to the wetlands of the study area (38.35 km\textsuperscript{2}). We transformed nest density to abundance of breeding individuals by multiplying it by a factor of two (assuming two individuals per nest). We acknowledge that the opportunistic monitoring of these species likely underestimated their true nest density. However, considering the extensive sampling effort deployed annually within this area, we are confident to obtain a realistic order of magnitude for the number of nests present. Because duck sightings are frequent, yet only a few nests are found, we believe there may be non-breeding individuals. Therefore, we employed an additional method to estimate the overall duck populations without differentiating between breeding and non-breeding individuals. 
As an alternative approach, we estimated the abundance of ducks based on the indices of relative abundance (i.e., the number of individuals observed per 100 hours) presented by \citet{gauthier2024a}. We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) in duck and loon species. We therefore derived the absolute abundance of long-tailed ducks and king eiders from their relative abundances using the ratio between relative and absolute abundances of red-throated loons as a reference.

@@

                \item[] \textit{\textbf{Sandhill crane}}\\
                We estimated the mean abundance of sandhill cranes in the lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:sacr}}). In this relationship, nest density and transect observations come from the 8 km\textsuperscript{2} area of the Qarlikturvik valley where extensive nest search is performed. We acknowledge that the opportunistic monitoring of this species likely underestimated the true nest density. However, considering the extensive sampling effort deployed annually within this area, we are confident in obtaining a realistic order of magnitude for the number of nests present. Number of individuals observed along transects in each lowland zone was converted into nest density using the regressions, and then in total number of individuals in each zone by multiplying by the area of the zone and a factor 2. We estimated the density in the upland zones by applying a correction factor to the annual mean density in lowland zones. This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).

\begin{figure}[H]
\centering
\includegraphics[width=0.5\linewidth]{figures/sandhill_crane_nest_transects.pdf} 
\caption{Linear regression between the nest density of sandhill cranes and the number of individuals observed per transect (nest density = 1.12 x number of individuals per transect; regression was forced to pass through the origin). The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots). Data points represent annual values.}
\label{figure:sacr}
\end{figure}
                \item[] \textit{\textbf{American golden-plover and black-bellied plover}}\\
                We applied a distance sampling approach to the observations of American golden-plovers made along vertebrate count transects to estimate the abundance of individuals in the mesic habitat of the study area between 2014 and 2023. We used perpendicular distance between detected individuals and transect (n= 1015) to estimate a detection function with the \textit{ds} function from the \textit{Distance} package \citep{miller2019}. We applied a truncation distance of 150 m (i.e., maximum distance on either side of the observer where observations have been considered). We selected the model with the lowest AIC, which included a "hn" key and a single "cos" adjustment term. We excluded observations of more than four individuals, as these likely indicated groups of non-breeders passing through the area. We did not estimate abundance in wetland habitat because American golden-plovers nest almost exclusively in mesic habitat \citep{parmelee1967}. We estimated the abundance in the upland zones (i.e., plateaus) by applying a correction factor to the abundance in lowland zones. This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).

To determine the abundance of black-bellied plovers, we used the mean number of black-bellied plovers and American golden-plovers observed per transect as an index of relative abundance. We assumed that the ratios of between relative and actual abundance are the same (i.e., similar detection probability) among those species. This assumption is realistic as those species present similarities in size, color, and reproductive behavior. We therefore derived the absolute abundance of black-bellied plovers from their relative abundance using the ratio between relative and absolute abundances of American golden-plover as a reference.
As an alternative approach to determine black-bellied plover abundance, we used the same approach as previously described, but with the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations.

@@

                \item[] \textit{\textbf{Lapland longspur and Baird’s sandpiper}}\\
                We estimated the mean abundance of Lapland longspur in the different lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:lalo_basa}}). For Baird’s sandpiper, we employed a similar approach, but instead of using the mean number of individuals observed per transect, we used the mean proportion of transects where at least one individual was detected. We made this adjustment because this species was less frequently observed. In this relationship, nest density for these two species came from the intensive nest sampling conducted within the core 2 km\textsuperscript{2} area of the Qarlikturvik valley and observations of individuals from transects carried out in the larger 8 km\textsuperscript{2} area in which the core area was located. This approach allowed us to incorporate a larger sample size from the transects while focusing on a measure of nest density determined systematically. Transects observations in lowland were then converted into nest density using the regressions, and then in total number of individuals by multiplying by the area and a factor 2. We estimated the density of both species in the upland zones by applying a correction factor to the annual mean density in lowland zones. This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). We acknowledge that the regression for Baird’s sandpiper is weak; however, it offers some refinement compared to assuming a uniform density throughout the study area.

@@

                \item[] \textit{\textbf{Other passerines and sandpipers}}\\
                We estimated the abundance of other passerines (horned lark, American pipit, and snow bunting) in the lowland zones of the study area with the regression equation between number of individuals per transect and nest density of the Lapland longspur (see section \textit{Lapland longspur and Baird’s sandpiper}). We assumed here a similar detection probability for all species. We used the same approach for other sandpiper species (white-rumped sandpiper, pectoral sandpiper, buff-breasted sandpiper, red knot, ruddy turnstone and red phalarope) based on the regression equation for the Baird's sandpiper (see section \textit{Lapland longspur and Baird’s sandpiper}). For all these species, we estimated the density in the upland zones by applying a correction factor to the mean density in lowland zones. This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). Nest density was then converted in number of individuals by multiplying by the area and a factor 2.
As an alternative approach, we estimated the abundance of other passerines and sandpipers based on the indices of relative abundance (i.e., the number of individuals observed per 100 hours) presented by \citet{gauthier2024a}. We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among both passerine and sandpiper species. We therefore derived the absolute abundance of other passerine and sandpiper species from their relative abundances using respectively, the ratios between relative and absolute abundances of Lapland longspur (passerines) and Baird’s sandpiper (sandpipers) as references.

@@

                \item[] \textit{\textbf{Parasitic jaeger}}\\
                Based on the opportunistic nest monitoring of parasitic jaegers across the study area, an average of 3.75 nests is found annually, a small number considering that parasitic jaegers were frequently observed at the study site \citep{gauthier2024a}. This suggests that there may be non-breeding individuals present at the study site, or alternatively, individuals may regularly travel long distances, potentially from outside the study area, to forage during the breeding season. Due to limited data availability for estimating the abundance of non-breeding parasitic jaegers, we relied on the maximum number of adults banded during a single year (17 individuals in 2009; Therrien, unpublished data) as the minimum abundance on the study area. This corresponds to a density of 0.04 individuals/km\textsuperscript{2}. For comparison, \citet{taylor1974} measured a density of 0.06 individual/km\textsuperscript{2} on Bathurst Island.

                \item[] \textit{\textbf{Common raven}}\\
                Despite the intensive nest searches for raptors in upland zones, we never found more than one common raven nest each year, a small number considering the frequent raven observations at the study site \citep{gauthier2024a}. This indicates the potential presence of non-breeding individuals or individuals that breed outside the study area but use it for foraging. Therefore, we opted for alternative approaches based on individual counts to estimate the abundance of both breeding and non-breeding ravens. 
As a first approach, we based our estimate of ravens on the number of glaucous gulls observed per transect. We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among those species. This assumption is reasonable as those species present similarities in size and foraging strategy. We therefore derived the absolute abundance of common ravens from their relative abundance using the ratio between relative and absolute abundances of glaucous gulls as a reference.  
Independently, we estimated the abundance of common ravens with the same approach but using the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations, rather than observations from the transects.

        \item[] \textit{\textbf{Brown and collared lemming}}\\
        Between 1995 and 2003, we used the density estimates derived from the snap-trapping indices obtained in late July in each habitat. Between 2004 and 2007, annual abundance of each lemming species was based on the late-July density estimates on trapping grid in wet and mesic habitats. However, starting from 2008, estimates were derived from the mean density recorded in mid-July and mid-August, except for two instances: 2019 and 2021. In 2019, due to an exceptionally early snowmelt and thus an early decline in lemmings during the summer, we only retained value from mid-July. In 2021, we relied solely on data gathered in August because it was the only trapping period carried out that year. To scale the estimated densities from the wet and mesic grids to the entire study area, we used the proportions of mesic habitats (64\%) and wet habitats (10\%) measured within the study area.

@@

                \item[] \textit{\textbf{Ermine}}\\
                We estimated the annual abundance of ermines by transforming the annual index of relative abundance provided in \citet{bolduc2023} into individual density. Annual values ranged from 0, indicating no ermine sighting, to 2.88, which signifies that nearly all observers observed at least one family group during their field season. We independtly obtained measures of minimum (0.02 ind./km\textsuperscript{2}) and maximum (0.4 ind./km\textsuperscript{2}) ermine density, which were determined from estimates of individual home range obtained from radio-tracking data, observations on Bylot Island, and existing literature \citep{legagneux2012, bilodeau2013}. We associated the minimum and maximum scores of relative abundance with the minimum and maximum density of individuals, respectively. Ultimately, we calculated the ermine density by linearly interpolating between these two density extremes using the annual index of relative abundance.

@@

                        \newgeometry{top=0.5cm, bottom=1.5cm, left=0.5cm, right=0.5cm}
\begin{landscape}
\input{tables/table_summary_methods_and_abundance.tex}
%label-> table:table_summary_methods_and_abundance
\end{landscape}
\restoregeometry
\newpage
\begin{figure}[h]
\caption{Time series of the estimated annual abundance of vertebrate species on the southern plain of Bylot Island (389 km\textsuperscript{2}). Estimated abundance represents adult individuals, with the exception of lemmings, for which juveniles were also included in the estimate. Time series shorter than 5 years are not presented.}
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Cackling_goose.pdf}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Snow_goose.pdf}
  \end{subfigure} 
    \hfill
    \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Rough-legged_hawk.pdf}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Peregrine_falcon.pdf}
  \end{subfigure}
\end{figure}
\newpage
\begin{figure}[h]
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Snowy_owl.pdf}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/American_golden-plover.pdf}
  \end{subfigure} 
    \hfill
    \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Glaucous_gull.pdf}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Long-tailed_jaeger.pdf}
  \end{subfigure}
\end{figure}
\newpage
\begin{figure}[h]
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Nearctic_brown_lemming.pdf}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Nearctic_collared_lemming.pdf}
  \end{subfigure} 
    \hfill
    \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Ermine.pdf}
  \end{subfigure}
\end{figure}
\newpage
\input{tables/table_expert_based_uncertainty.tex}
%label-> table:table_expert_based_uncertainty
\newpage

>>>>

\begingroup  
  \centering
\textbf{Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut}\\[1.5em]
 Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Gilles Gauthier\textsuperscript{4}, Éliane Duchesne\textsuperscript{1}, 
 Dominique Fauteux\textsuperscript{5}, Dominique Berteaux\textsuperscript{1}, Pierre Legagneux\textsuperscript{3,6}, Marie-Christine Cadieux \textsuperscript{3} and Joël Bêty\textsuperscript{1}\\[1.5em]
\textbf{Author Affiliations}:\\
\textsuperscript{1} Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada\\
\textsuperscript{2} Chaire de Recherche du Canada en Écologie Intégrative, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université de Sherbrooke, Sherbrooke, QC, Canada\\
\textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’Impact des Migrations Animales sur les Écosystèmes Nordiques, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{4} Centre d’Études Nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{5} Centre d’Études Nordiques, Centre de Connaissance et d’Exploration de l’Arctique, Musée Canadien de la Nature, Ottawa, ON, Canada\\
\textsuperscript{6} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France 
\\[1.5em]
\textbf{Corresponding Author}:\\
\textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
\textit{\href{mailto:joel_bety@uqar.ca}{joel{\textunderscore}bety@uqar.ca}}
\\[1.5em]
\textbf{Open Research statement}:\\
The data set is publicly available at \url{https://datadryad.org/XXXXXXX} and the raw data and codes used to extract the data set are publicly available at \url{https://zenodo.org/XXXXXXXXXX}.\\
\endgroup
\newpage

@@

Since 2004, Nearctic brown and collared lemmings were live-trapped 3 times during the summer (mid-June, mid-July, and mid-August) in two 11 ha grids. Each grid is made of 144 traps separated by 30 m according to a cartesian plane, one in mesic habitat and the other in wet habitat, located in the Qarlikturvik valley \citep{fauteux2015, gauthier2020lemmings}. Density of each species was estimated at each occasion using spatially explicit capture-recapture methods (see \citet{fauteux2015} for details). From 1995 to 2016 snap-trapping was performed once a year (mid-July) along 2 groups of transects located in the same habitats than the trapping grids \citep{gruyer2008}. Index of abundance derived from snap-trapping were transformed in density estimates in each habitat for the period 1995-2003 using the equation provided by \citet{fauteux2018} based on the period of overlap between the two sampling methods (2004 to 2016). 

@@

All vertebrate individuals captured for marking purposes were systematically weighed: snow goose (G. Gauthier, M.-C. Cadieux and J. Lefebvre, unpublished data), snowy owl \citep{therrien2012, robillard2018}, American-golden plovers \citep{lamarre2021}, common-ringed plovers \citep{leandri2019}, other shorebirds (J. Bêty, unpublished data), glaucous gulls \citep{gauthier2015}, long-tailed jaeger \citep{seyer2019}, parasitic jaegers (J.-F. Therrien and G. Gauthier, unpublished data), Lapland longspurs (J. Bêty and G. Gauthier, unpublished data), lemmings \citep{gauthier2020lemmings}, American ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}. When not available, we extracted mean body mass from the literature \citep{wilman2014}.

@@

    \subsection*{3.  Research methods}
                \subsubsection*{a. Field/laboratory}
We estimated the abundance of breeding individuals for most species, but there were a few exceptions. For common ravens, parasitic jaegers, long-tailed ducks, and king eiders, we suspect the presence of a significant number of non-breeding individuals in the study area. Therefore, the estimates we provided for these species include both breeding and potentially non-breeding individuals. Additionally, we did not distinguish between breeding and non-breeding individuals for mammals such as brown and collared lemmings, Arctic fox, American ermine, and Arctic hare. The methods used for each species are summarized in (\textbf{Table \ref{table:table_summary_methods_and_abundance}}).

@@

                \item[] \textit{\textbf{King eider and long-tailed duck}}\\
                We first estimated the abundance of both king eiders and long-tailed ducks based on the annual nest density of each species found in the 8 km\textsuperscript{2} extensive nest search area located in the Qarlikturvik valley. We extrapolated the mean nest density in the wetlands of the Qarlikturvik valley to the wetlands of the study area (38.35 km\textsuperscript{2}). We transformed nest density to abundance of breeding individuals by multiplying it by a factor of two (assuming two individuals per nest). We acknowledge that the opportunistic monitoring of these species likely underestimated their true nest density. However, considering the extensive sampling effort deployed annually within this area, we are confident to obtain a realistic order of magnitude for the number of nests present. Because duck sightings are frequent throughout the breeding period, yet only a few nests are found, we believe there may be a significant portion of non-breeding individuals. Therefore, we employed an additional method to estimate the overall duck populations without differentiating between breeding and non-breeding individuals. 
As an alternative approach, we estimated the abundance of ducks based on the indices of relative abundance (i.e., the number of individuals observed per 100 hours) presented by \citet{gauthier2024a}. We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) in duck and loon species. We therefore derived the absolute abundance of long-tailed ducks and king eiders from their relative abundances using the ratio between relative and absolute abundances of red-throated loons as a reference.

@@

                \item[] \textit{\textbf{Sandhill crane}}\\
                We estimated the mean abundance of sandhill cranes in the lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:sacr}}). In this relationship, nest density and transect observations come from the 8 km\textsuperscript{2} area of the Qarlikturvik valley where extensive nest search is performed. We acknowledge that the opportunistic monitoring of this species likely underestimated the true nest density. However, considering the extensive sampling effort deployed annually within this area, we are confident in obtaining a realistic order of magnitude for the number of nests present. Number of individuals observed along transects in each lowland zone was converted into nest density using the regressions, and then in total number of individuals in each zone by multiplying by the area of the zone and a factor 2. We estimated the density in the upland zones by applying a correction factor to the annual mean density in lowland zones. This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).

\begin{figure}[H]
\centering
\includegraphics[width=0.5\linewidth]{figures/sandhill_crane_nest_transects.pdf} 
\caption{Linear regression between the nest density of sandhill cranes and the number of individuals observed per transect (nest density = 1.12 x number of individuals per transect; regression was forced to pass through the origin). The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots). Data points represent annual values.}
\label{figure:sacr}
\end{figure}
                \item[] \textit{\textbf{American golden-plover and black-bellied plover}}\\
                We used a distance sampling approach to estimate the abundance of American golden-plovers in the lowland zones of the study area between 2014 and 2023. Observations of plovers were made along vertebrate count transects mainly in mesic habitat. Perpendicular distance between detected individuals and the transect path were used (n= 1015) to estimate a detection function with the \textit{ds} function from the \textit{Distance} package \citep{miller2019}. To determine the detection function, we applied a truncation distance of 150 m (i.e., maximum distance on either side of the observer where observations have been considered) and selected the model with the lowest AIC, which included a "hn" key and a single "cos" adjustment term. We excluded observations of groups with more than four individuals, as these likely indicated groups of non-breeders. We did not estimate abundance in wetland habitat because American golden-plovers nest almost exclusively in mesic habitat \citep{parmelee1967}. We estimated the abundance in the upland zones (i.e., plateaus) by applying a correction factor to the abundance in lowland zones. This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).

To determine the abundance of black-bellied plovers, we used the mean number of black-bellied plovers and American golden-plovers observed per transect as an index of relative abundance. We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among those species. This assumption is realistic as those species present similarities in size, color, and reproductive behavior. We therefore derived the absolute abundance of black-bellied plovers from their relative abundance using the ratio between relative and absolute abundances of American golden-plover as a reference. As an alternative approach to determine black-bellied plover abundance, we used the same approach as previously described, but with the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations.

@@

                \item[] \textit{\textbf{Lapland longspur and Baird’s sandpiper}}\\
                We estimated the mean abundance of Lapland longspur in the different lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:lalo_basa}}). For Baird’s sandpiper, we employed a similar approach, but instead of using the mean number of individuals observed per transect, we used the mean proportion of transects where at least one individual was detected. We made this adjustment because this species was less frequently observed. In this relationship, nest density for these two species came from the intensive nest sampling conducted within the core 2 km\textsuperscript{2} area of the Qarlikturvik valley and observations of individuals from transects carried out in the larger 8 km\textsuperscript{2} area in which the core area was located. This approach allowed us to incorporate a larger sample size from the transects while focusing on a measure of nest density determined systematically. Transects observations in lowland were then converted into nest density using the regressions, and then in total number of individuals by multiplying by the area and a factor 2. We estimated the density of both species in the upland zones by applying a correction factor to the annual mean density in lowland zones. This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). We acknowledge that the regression for Baird’s sandpiper is weak; however, it offers some refinement compared to assuming a uniform density throughout the study area.

@@

                \item[] \textit{\textbf{Other passerines and sandpipers}}\\
                We estimated the abundance of other passerines (horned lark, American pipit, and snow bunting) in the lowland zones of the study area with the regression equation between number of individuals per transect and nest density of the Lapland longspur (see section \textit{Lapland longspur and Baird’s sandpiper}). We assumed here a similar detection probability for all species. We used the same approach for other sandpiper species (white-rumped sandpiper, pectoral sandpiper, buff-breasted sandpiper, red knot, ruddy turnstone and red phalarope) based on the regression equation for the Baird's sandpiper (see section \textit{Lapland longspur and Baird’s sandpiper}). For all these species, we estimated the density in the upland zones by applying a correction factor to the mean density in lowland zones. This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). Nest density was then converted in number of individuals by multiplying by the area and a factor 2.
As an alternative approach, we estimated the abundance of other passerines and sandpipers based on the indices of relative abundance (i.e., the number of individuals observed per 100 hours) presented by \citet{gauthier2024a}. We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among both passerine and sandpiper species. We therefore derived the absolute abundance of other passerine and sandpiper species from their relative abundances using respectively, the ratios between relative and absolute abundances of Lapland longspur (passerines) and Baird’s sandpiper (sandpipers) as references.

@@

                \item[] \textit{\textbf{Parasitic jaeger}}\\
                Based on the opportunistic nest monitoring of parasitic jaegers across the study area, an average of 4 nests is found annually, a small number considering that parasitic jaegers were frequently observed at the study site \citep{gauthier2024a}. This suggests that there may be non-breeding individuals present at the study site, or alternatively, individuals may regularly travel long distances, potentially from outside the study area, to forage during the breeding season. Due to limited data availability for estimating the abundance of non-breeding parasitic jaegers, we relied on the maximum number of adults banded during a single year (17 individuals in 2009; Therrien, unpublished data) as the minimum abundance on the study area. This corresponds to a density of 0.04 individuals/km\textsuperscript{2}. For comparison, \citet{taylor1974} measured a density of 0.06 individual/km\textsuperscript{2} on Bathurst Island.

                \item[] \textit{\textbf{Common raven}}\\
                Despite the intensive nest searches for raptors in upland zones, we never found more than one common raven nest each year, a small number considering the frequent raven observations at the study site \citep{gauthier2024a}. This indicates the potential presence of non-breeding individuals or individuals that breed outside the study area but use it for foraging throughout the breeding period. Therefore, we opted for alternative approaches based on individual counts to estimate the abundance of both breeding and non-breeding ravens. 
As a first approach, we based our estimate of ravens on the number of glaucous gulls observed per transect. We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among those species. This assumption is reasonable as those species present similarities in size and foraging strategy. We therefore derived the absolute abundance of common ravens from their relative abundance using the ratio between relative and absolute abundances of glaucous gulls as a reference.  
Independently, we estimated the abundance of common ravens with the same approach but using the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations, rather than observations from the transects.

        \item[] \textit{\textbf{Nearctic brown and collared lemming}}\\
        Between 1995 and 2003, we used the density estimates derived from the snap-trapping indices obtained in late July in each habitat. Between 2004 and 2007, annual abundance of each lemming species was based on the late-July density estimates on trapping grid in wet and mesic habitats. However, starting from 2008, estimates were derived from the mean density recorded in mid-July and mid-August, except for two instances: 2019 and 2021. In 2019, due to an exceptionally early snowmelt and thus an early decline in lemmings during the summer, we only retained value from mid-July. In 2021, we relied solely on data gathered in August because it was the only trapping period carried out that year. To scale the estimated densities from the wet and mesic grids to the entire study area, we used the proportions of mesic habitats (64\%) and wet habitats (10\%) measured within the study area.

@@

                \item[] \textit{\textbf{American ermine}}\\
   				We estimated the annual abundance of ermines by transforming the annual index of relative abundance provided in \citet{bolduc2023} into individual density. Annual values ranged from 0, indicating no ermine sighting, to 2.88, which signifies that nearly all observers observed at least one family group during their field season. We independtly obtained measures of minimum (0.02 ind./km\textsuperscript{2}) and maximum (0.4 ind./km\textsuperscript{2}) ermine density, which were determined from estimates of individual home range obtained from radio-tracking data, observations on Bylot Island, and existing literature \citep{legagneux2012, bilodeau2013}. We associated the minimum and maximum scores of relative abundance with the minimum and maximum density of individuals, respectively. Ultimately, we calculated the ermine density by linearly interpolating between these two density extremes using the annual index of relative abundance.

@@

                        \newgeometry{top=0.5cm, bottom=1.5cm, left=0.5cm, right=0.5cm}
\begin{landscape}
\input{tables/table_summary_methods_and_abundance.tex}
%label-> table:table_summary_methods_and_abundance
\end{landscape}
\restoregeometry
\newpage
\begin{figure}[h]
\caption{Time series of the estimated annual abundance of vertebrate species on the southern plain of Bylot Island (389 km\textsuperscript{2}). Estimated abundance represents adult individuals, with the exception of lemmings, for which juveniles were also included in the estimate. Time series shorter than 5 years are not presented.}
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Cackling_goose.pdf}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Snow_goose.pdf}
  \end{subfigure} 
    \hfill
    \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Rough-legged_hawk.pdf}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Peregrine_falcon.pdf}
  \end{subfigure}
\end{figure}
\newpage
\begin{figure}[h]
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Snowy_owl.pdf}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/American_golden-plover.pdf}
  \end{subfigure} 
    \hfill
    \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Glaucous_gull.pdf}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Long-tailed_jaeger.pdf}
  \end{subfigure}
\end{figure}
\newpage
\begin{figure}[h]
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Nearctic_brown_lemming.pdf}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/Nearctic_collared_lemming.pdf}
  \end{subfigure} 
    \hfill
    \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/American_ermine.pdf}
  \end{subfigure}
\end{figure}
\newpage
\input{tables/table_expert_based_uncertainty.tex}
%label-> table:table_expert_based_uncertainty
\newpage
